# NYC Bike Hire Analysis 
## By Graham Angus 

**Introduction**


```{r}
library(tsibbledata)
library(tidyverse)
library(lubridate)
library(leaflet)
library(slider)
library(scales)

nyc <- tsibbledata::nyc_bikes
nyc_bikes_df <- nyc

nyc_bikes_sum <-
  nyc_bikes_df %>%
  mutate(year = year(start_time)) %>%
  mutate(month = month(start_time)) %>%
  mutate(day = day(start_time)) %>%
  group_by(month) %>%
  summarise(hire_count = n())

nyc_bikes_df <-
  nyc_bikes_df %>% mutate(nyc_bikes_df$start_time, hour_start = format(nyc_bikes_df$start_time, format = "%H"))

nyc_bikes_plot <-
  nyc_bikes_df %>%
  mutate(year = year(start_time)) %>%
  mutate(month = month(start_time)) %>%
  mutate(day = day(start_time)) %>%
  group_by(hour_start) %>%
  summarise(hire_count = n())

nyc_gender <- nyc_bikes_df %>% 
  group_by(gender, type) %>%
  summarise(count = n())

nyc_bikes_df$diff_in_mins <- difftime(nyc_bikes_df$stop_time, nyc_bikes_df$start_time, units = c("mins"))

head_bikes <-
nyc_bikes_df %>%
head(nyc_bikes_df, n = 40)

nyc_cycle_map <- leaflet() %>%
  addTiles() %>%  
  addMarkers(lng = head_bikes$start_long, lat = head_bikes$start_lat)

nyc_bikes_df %>%
  ggplot() +
  aes(x = birth_year, y = diff_in_mins, colour = gender) +
  geom_point() +
  scale_x_continuous(limits = c(1945, 2000)) +
  scale_y_continuous(limits = c(1, 75)) +
  ylab("Duration") +
  xlab("Birth Year") +
  facet_wrap(~ gender)    



knitr::opts_chunk$set(echo = FALSE)

```


```{r}


nyc <- tsibbledata::nyc_bikes
nyc_bikes_df <- nyc
```

**The Gender Divide **


```{r}
# Citi Bike stated goals are Save money, Save Time, Have Fun, Get Exercise, Go green

ggplot(nyc_gender) +
  geom_col(aes(x = gender, y = count, fill = type))

# 23% of users are female

nyc_bikes_df %>%
  group_by(gender, type) %>%
  summarise(mean_duration = mean(diff_in_mins))

# The mean duration for females is 14mins, for men it is 7mins

nyc_bikes_df %>%
  ggplot() +
  aes(x = birth_year, y = diff_in_mins, colour = gender) +
  geom_point() +
  scale_x_continuous(limits = c(1945, 2000)) +
  scale_y_continuous(limits = c(1, 75)) +
  ylab("Duration") +
  xlab("Birth Year") +
  facet_wrap(~ gender)  

# Age skewed to lower limit

nyc_bikes_df %>%
  group_by(gender) %>%
summarise(mode_age = median(birth_year))

# Median age for females is 34, 33 for men

library(scales)
nyc_bikes_df %>%
ggplot() +
  aes(x = hour_start, colour = gender) +
  geom_bar(position = "dodge") 

# 8am and 6pm are the two peak hours for both genders.

nyc_cycle_map <- leaflet() %>%
  addTiles() %>%  
  addMarkers(lng = head_bikes$start_long, lat = head_bikes$start_lat)
nyc_cycle_map

# Most popular areas for pick up

```

# Conclusion

There is a big gender divide in the companies users and with further inspection I 
have found that the 4 main reasons for woman not cycling are Appearance, 
traffic/intimidation, their age, lack of fitness. 

I suggest that a good way to meet the aims of the company would be to try to 
engage female users via solutions to incentivise off peak cycling, potentially 
looking at learn to ride sessions and off peak memberships for those of a certain
age and female friendly seats.

If they really want to meet these aims then they could do worse than offering 
support to younger females as 79% of NYC under 16s know how to ride a bike compared
to 91% of males, fixing this end of the pipeline with some early community engagement 
could be huge for the company.

Only 1.1% of NYC residents regularly commute via bike, the duration figures and 
the travel patterns suggest this is the main demographic of Citi Bike, there is 
a potenially lucrative market from Tourist users which again could fill up the 
off peak timetable. 