---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

rev_breakdown <- CodeClanData::qb_revenue_breakdown
month_sales <- CodeClanData::qb_monthly_sales
competitors <- CodeClanData::qb_competitors
device_data <- CodeClanData::qb_device_data

```

# Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(rev_breakdown) +
  geom_col() +
  aes(x = Product, y = Revenue, fill = Year) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  facet_wrap(~ Year)
```

# Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
flagship_product <- 
  filter(month_sales, 
  Cashflow == "Kwikbit Unit 1 Sales")

flagship_product

ggplot(flagship_product) +
  geom_line(aes(x = Date, y = Amount))
```

# Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
cost_revenue_monthly <- 
  filter(month_sales,
         Cashflow == "Total Revenue" | Cashflow == "Cost")

ggplot(cost_revenue_monthly) +
  geom_line(aes(x = Date, y = Amount, colour = Cashflow))
  
```

# Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
ggplot(competitors) +
  geom_area() +
  aes(x = Year, y = Revenue) +
  facet_wrap(~ Company)
```

# Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(competitors) +
  geom_line() +
  aes(x = Year, y = Revenue, colour = Company) +
  facet_wrap(~ Company)
```

# Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
ggplot(device_data) +
  geom_line() +
  aes(x = c(hours, mins), y = counts)
  
```


