---
title: "R programming - weekend homework - answers"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_cat: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Learning Objectives

* A chance to practice everything you've learnt over this week.
* Practice using `if` statements, `for` loops and `while` loops.
* Practice writing your own functions.
* Be able to solve puzzles, using code.
* Be able to help yourself when you are stuck and debug problems.
* Create a cheatsheet of useful base R functions and syntax as you go.

# MVP

```{r, echo = FALSE, eval = TRUE}
# Just a helper function to make the tables seen in the questions. Not part of the answers
make_table <- function(fun, input) {
  output <- purrr::map(input, fun)

  table <- data.frame(
    Input = unlist(input),
    Output = unlist(output)
  )

  knitr::kable(
    table
  )
}
```

## Cheatsheet

We'd like you to create your own cheatsheet of base R functionality as you go through these exercises, note down useful functions and syntax as you use them. It doesn't need to be pretty, functional is good enough!

## `if` statements and functions

For each of these questions, write a function. After each question we'll give you some example input and output. Check your function gives the expected output for each input.

###

Write a function which multiples all numbers by 15 and adds 4.

```{r, eval = TRUE}
times_15_add_4 <- function(x) {
  x * 15 + 4
}
```

```{r, echo = FALSE, eval = TRUE}
make_table(times_15_add_4, c(1, 2, 3))
```

###

Write a function that returns "bird", if the input is "goose", "duck" or "pigeon", and returns "not bird" for all other input.

```{r, eval = TRUE}
is_bird <- function(x) {
  if (x %in% c("goose", "duck", "pigeon")) {
    return("bird")
  } else {
    return("not bird")
  }
}
```

```{r, echo = FALSE, eval = TRUE}
make_table(is_bird, c("goose", "apple", 5, "seagull"))
```

###

Write a function which multiplies all numbers by 10. If the input isn't a number, the function should return "Invalid input".

```{r, eval = TRUE}
times_10 <- function(x) {
  if (is.numeric(x)) {
    return(x * 10)
  } else {
    return("Invalid input")
  }
}
```


```{r, echo = FALSE, eval = TRUE}
make_table(times_10, list(2, 5.3, "x", 7L))
```


## `ifelse()`, `recode()` and `case_when()`

For these questions, we want you to code solutions *without* using `if` statements or loops. 

Each question can be answered with either `ifelse()`, `recode()` or `case_when()`. Decide which function you think works best and use it to answer the question.

Remember to load `dplyr` to use `recode()` and `case_when()`.

```{r, message = FALSE, warning = FALSE}
library(dplyr)
```

###

For the vector below, make a new vector which is equal to "Big" every time the element of `sizes` is greater than 10, and equal to "Small" every time the element of `sizes` is less than or equal to 10.

```{r, echo = TRUE}
sizes <- c(13, 15, 10, 8, 22)
```


```{r}
# ifelse
ifelse(sizes > 10, "Big", "Small")

# recode
recode(sizes, `10` = "Small", `8` = "Small", .default = "Big")

# case_when
case_when(
  sizes > 10 ~ "Big",
  TRUE ~ "Small"
)

# ifelse is the best choice here. case_when is also okay
```

### 

For the vector below convert:

apple → alpaca

banana → bonobo

cherry → chipmunk


```{r, echo = TRUE}
fruit <- c("apple", "apple", "banana", "cherry", "cherry")
```


```{r}
# if else
animals <- ifelse(fruit == "apple", "alpaca", fruit)
animals <- ifelse(animals == "banana", "bonobo", animals)
ifelse(animals == "cherry", "chipmunk", animals)

# recode
recode(fruit, apple = "alpaca", banana = "bonobo", cherry = "chipmunk")

# case_when
case_when(
  fruit == "apple" ~ "alpaca",
  fruit == "banana" ~ "bonobo",
  fruit == "cherry" ~ "chipmunk"
)

# recode is easiest here. Again, case_when is an okay choice.
```

###

The two vectors below show the ages at marriage of each member of a couple. The first elements of each vector correspond to the first couple, the second elements to the second couple, and so on.

Create a new vector which is equal to:

"Wife older" whenever the wife's age is highest
"Husband older" when the husband's age is highest
"Same age" when they are the same age

```{r, echo = TRUE, eval = TRUE}
wife_age <- c(29L, 30L, 36L, 31L, 22L)
husband_age <- c(28L, 35L, 32L, 31L, 29L)
```


```{r}
# ifelse
ifelse(wife_age > husband_age, "Wife older", ifelse(wife_age < husband_age, "Husband older", "Same age"))

# recode
difference <- sign(wife_age - husband_age)
recode(difference, `1` = "Wife older", `0` = "Same age", `-1` = "Husband older")

# case_when
case_when(
  wife_age > husband_age ~ "Wife older",
  wife_age < husband_age ~ "Husband older",
  wife_age == husband_age ~ "Same age"
)

# case_when is easiest to write, and easiest to understand
```

## Loops

For each question on loops we're going to use the list below:

```{r, echo = TRUE}
my_list <- list(
  4,
  "apple",
  10.5,
  c("is", "an", "apple"),
  "apple",
  c("i", "like", "apples"),
  8
)
```


###

Use a loop to print out every element of `my_list` which is of type `character`. Decide if you need a `for` loop or a `while` loop.

```{r}
for (elem in my_list) {
  if (is.character(elem)) {
    print(elem)
  }
}
```

###

Once again write a loop that prints out character elements. This time if the element has a `length()` greater than one, `paste()` all the elements in that vector together into one big character vector.

```{r}
for (elem in my_list) {
  if (is.character(elem)) {
    if (length(elem) > 1) {
      print(paste(elem, collapse = " "))
    } else {
      print(elem)
    }
  }
}
```

# Extension - Puzzles

##

Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

*From: https://adriann.github.io/programming_problems.html*

```{r}
matrix(1:12) %*% t(matrix(1:12))
```


##

Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function `cat()`. This is similar to `print()`, but prints directly, without vector markers. You can also include newlines `"\n"` when using `cat()`.

```
1 | ***
2 | *
3 | *******
4 | ****
5 | **
```

```{r}
make_histogram <- function(x){
  
  distinct <- sort(unique(x))
  
  for (i in distinct){
    
    count <- sum(x == i)
    
    cat(i)
    cat(" | ")
    cat(rep("*", count))
    cat("\n")
    
  }
  
}
```

##

Write a guessing game where the user has to guess a secret number. After every guess the program tells the user whether their number was too large or too small. At the end the number of tries needed should be printed. 

To take user input, use the function `readline()`. Whenever R encounters `readline`, the code will stop, and the value assigned from `readline` will depend on the value that the user inputs. For more information see the help file.

*From: https://adriann.github.io/programming_problems.html*

```{r, eval = FALSE}
number <- round(runif(1, 1, 10))

guessed <- FALSE
attempts <- 0

while (!guessed){
  
  print("Guess a number between 1 and 10!")
  
  guess <- readline("")
  
  if (guess > number){
    
    print("Guess is too high")
    
    attempts <- attemps + 1
    
  }
  
  if (guess < number){
    
    print("Guess too low")
    
    attempts <- attempts + 1
    
  }
  
  if (guess == number){
    
    print("Guess is right! Well done!")
    
    print(paste("Attempts:", attempts))
    
    guessed <- TRUE
    
  }
  
}
```

##

By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

You may want to read up on the definition of the Fibonacci sequence.

*From: https://projecteuler.net/problem=2*

```{r}
n_0 <- 0
n_1 <- 1

sum <- 0

while(n_1 <= 4000000){
  
  n_1 <- n_0 + n_1
  n_0 <- n_1 - n_0 # old n_1 value
  
  if (n_1 %% 2 == 0){
    sum <- sum + n_1
  }
  
}

sum
```

