---
title: "Decision Tree Homework"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(GGally)
library(janitor)
library(modelr)
library(yardstick)
library(caret)
titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

# Cleaning up the data is always the first step. Do the following:

    Take only observations which have a survived flag (i.e. that aren’t missing)
    Turn your important variables into factors (sex, survived, pclass, embarkation)
    Create an age_status variable which groups individuals under (and including) 16 years of age into a category called “child” category and those over 16 into a category called “adult”.
    Drop the NA
    Drop any variables you don’t need (X1, passenger_id, name, ticket, far, cabin)


```{r}
titanic_clean <- titanic_set %>%
  filter(survived %in% c(0,1)) %>%
    mutate(sex = as.factor(sex), 
           age_status = as.factor(if_else(age <= 16, "child", "adult")),
         class = factor(pclass, levels = c(3,2,1), labels = c("Lower", "Middle", "Upper")), 
           survived_flag = factor(survived, levels = c(0,1), labels = c("No", "Yes")), 
           port_embarkation = as.factor(embarked)) %>%
  select(sex, age_status, class, port_embarkation, sib_sp, parch, survived_flag) %>%
  na.omit()
```

Have a look at your data and create some plots to ensure you know what you’re working with before you begin. Write a summary of what you have found in your plots. Which variables do you think might be useful to predict whether or not people are going to die? Knowing this before you start is the best way to have a sanity check that your model is doing a good job.





```{r}
ggpairs(titanic_clean)
```

# it looks like there is potential predictor value from sex and port_embarkation

```{r}
model1 <- glm(survived_flag ~ sex, family = binomial(link = "logit"),
              data = titanic_clean)
summary(model1)


```

```{r}
n_data <- nrow(titanic_clean)

test_index <- sample(1:n_data, size = n_data*0.2)

titanic_test  <- slice(titanic_clean, test_index)

titanic_train <- slice(titanic_clean, -test_index)

titanic_test %>%
 janitor::tabyl(survived_flag)

```


```{r}
titanic_train %>%
 janitor::tabyl(survived_flag)
```

```{r}
titanic_fit <- rpart(
  formula = survived_flag ~ ., 
  data = titanic_train, 
  method = 'class'
)

rpart.plot(titanic_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 2)
```





```{r}
titanic_test_pred <- titanic_test %>%
  add_predictions(titanic_fit, type = 'class')
```

```{r}
titanic_test_pred %>%
  select(sex, class, sib_sp, survived_flag, pred)
```

```{r}
conf_mat <- titanic_test_pred %>%
              conf_mat(truth = survived_flag, estimate = pred)

conf_mat
```

```{r}
model_accuracy <- titanic_test_pred %>%
 accuracy(truth = survived_flag, estimate = pred)

model_accuracy 
```

```{r}
confusionMatrix(titanic_test_pred$pred, titanic_test_pred$survived_flag)
```

# My decision tree shows that Females from a lower class had a 0.95 probability 
of dying, females in general had a 0.76 probability of surviving, men had a 0.21 
probability. If you had siblings you had 0.30 probability of dying and if you didn't
then a 0.57 probability. This shows that women had a greater chance of surviving, if you
had siblings also improved your chances maybe as kids were a priority for getting onto
lifeboats, class also had a massive bearing on outcome.

My model accuracy is 0.7957746 which is statistically significant
